\large{\bf Looking forwards and backwards: dynamics and genealogies of locally regulated populations }

### 2023-12-15 review-responses.tex

\begin{point}{}
    Figure \ref{fig:super_vs_det_2d}: What is $N$, the time and other parameters (e.g. $\rho_r$ , $\rho_\gamma$) in the simulations? Same
for other figures. Please at least specify the range of values.
\end{point}

\reply{
    We've added some more detail to the caption.
}

\begin{point}{\revref}
% P12 paragraph after Corollary 2.11:
A more precise reference, or explanation for why
uniqueness holds for this special case would be helpful, as the references you ave are all
quite extensive.
\end{point}

\reply{
    \comment{TODO:
    Perhaps Evans/Perkins 1994 ``Measure-valued branching diffusions'' for Dawson's Girsanov thm?
    (reproduced as Thm IV.1.6 in Perkins, "Dawson-Watanabe Superprocesses...").}
    But: that is for two species.
}

@@

\begin{point}{}
P16: Second paragraph:
The notation $L_t$ seems to first appear here and was not yet defined.
\end{point}

\reply{
    Ah, true - now we introduce it (albeit still informally) earlier \llname{Linf_mention}.
}

\begin{point}{}
% P19-20:
    The equations for the lineage dynamics for the three examples [i.e., Fisher-KPP \revref, Allen-Cahn, and Porous Medium] are interesting.
Which is/are new in this paper, and which was/were known before (and related references)?
\end{point}

\reply{
    \comment{TODO:
    The PME is new;
    probably the others are in Birzu/Korolev/Hallaschek? 
    Also maybe see Roques et al 2008ish PNAS?
    It reduces to BM in a potential:
    find other references for this (i.e., the $\grad \log$...)
    }
}

@@

\begin{point}{}
This maybe out of the scope of this paper, but it is related. As the authors pointed out, two
lineages probably will not coalesce in the limiting model. However, they will coalesce in
the finite-$N$ model, though it may take a very long time. I wonder if the authors have a
feeling about whether the description of the lineage dynamics in the limit still be a good
approximation for that of the finite model for such a long time.
\end{point}

\reply{
    Good question -- we've added a note about this question \revref.
}

### 2023-12-15 forwards-and-backwards.tex
.

\begin{corollary} \label{cor:superprocess_uniqueness}
    Under the assumptions of Theorem~\ref{thm:nonlocal_convergence},
    if the martingale problem
    defined by equations~\eqref{eqn:limiting_mgale_problem} 
and~\eqref{eqn:limiting_mgale_variation}
    has a unique solution,
    then $(\eta^N_t)_{t \ge 0}$ converges weakly
to that solution
    as $N \to \infty$.
\end{corollary}
When $\alpha>0$, the limit points can be thought of as interacting superprocesses. For
example, when $r$ and $\gamma$ are constant, and 
$F(x,\eta_s)=1-\smooth{F} \eta_s(x)$,
we recover a superprocess with nonlinear death rates corresponding to logistic growth
\citep{etheridge:2004}
that is a continuous limit of the Bolker-Pacala model \citep{bolker/pacala:1997,bolker/pacala:1999}.
We are not aware of a general result to determine when we will have uniqueness of 
solutions to the martingale problem of Theorem~\ref{thm:nonlocal_convergence}
when $\alpha>0$.
However, the Dawson--Girsanov transform tells us that we have uniqueness \revpoint{1}{12}
in this special case of the superprocess with nonlinear death rates, and 
Perkins stochastic calculus (and its adaptation to a 
lookdown setting) provides uniqueness for cases with interactions in the 
dispersal mechanism of
the superprocess. We refer 
to~\cite{dawson:1993, perkins:1992}, and \cite{donnelly/kurtz:1999} 
for approaches to showing that these sorts of martingale problems are well-posed.

@@

We focus on the simplest situation, that in which the population process is deterministic.
However, the results here apply when the population process solves 
either a nonlocal or a classical PDE.
There will be no 
coalescence of ancestral lineages in the deterministic limit,
but understanding motion of single lineages is useful in practice,
and our results can
be seen as a first step towards understanding genealogies for high population densities.

>>>>

\begin{corollary} \label{cor:superprocess_uniqueness}
    Under the assumptions of Theorem~\ref{thm:nonlocal_convergence},
    if the martingale problem
    defined by equations~\eqref{eqn:limiting_mgale_problem} 
and~\eqref{eqn:limiting_mgale_variation}
    has a unique solution,
    then $(\eta^N_t)_{t \ge 0}$ converges weakly
to that solution
    as $N \to \infty$.
\end{corollary}
When $\alpha>0$, the limit points can be thought of as interacting superprocesses. For
example, when $r$ and $\gamma$ are constant, and 
$F(x,\eta_s)=1-\smooth{F} \eta_s(x)$,
we recover a superprocess with nonlinear death rates corresponding to logistic growth
\citep{etheridge:2004}
that is a continuous limit of the Bolker-Pacala model \citep{bolker/pacala:1997,bolker/pacala:1999}.
We are not aware of a general result to determine when we will have uniqueness of 
solutions to the martingale problem of Theorem~\ref{thm:nonlocal_convergence}
when $\alpha>0$.
However, the Dawson--Girsanov transform tells us that we have uniqueness \revpoint{1}{12}
in this special case of the superprocess with nonlinear death rates \cite{dawson:1993}.
The Perkins stochastic calculus
(and its adaptation to a lookdown setting \cite{donnelly/kurtz:1999})
provides uniqueness for a different but related class of processes,
in which interactions affect the dispersal mechanism
(rather than reproduction) of the superprocess \cite{perkins:1992}.
We refer to~\cite{dawson:1993, perkins:1992}, and \cite{donnelly/kurtz:1999} 
for approaches to showing that these sorts of martingale problems are well-posed.

@@

We focus on the simplest situation, that in which the population process is deterministic.
However, the results here apply when the population process solves 
either a nonlocal or a classical PDE.
There will be no coalescence of ancestral lineages in the deterministic limit,
but understanding motion of single lineages is useful in practice,
and our results can
be seen as a first step towards understanding genealogies for high population densities. 
Since the time scale on which coalescence occurs \revpoint{1}{17}
goes to infinity in the deterministic limit,
an important question to answer will be whether this description of lineage motion
is a good approximation over such a long time scale.
Other information may be important -- for instance, in \citet{etheridge/pennington:2022}
the form of the coalescent that is obtained depends
on fluctuations happening on a longer time scale than the mixing time of a lineage.
